Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-21 paper/main.tex
paper updates

We follow the overall strategy of \cite{learning-to-blame} and represent each
buggy program by a Bag of Abstracted Terms. In our setting, a `term' means either
a statment or an expression. For each term in the program we compute the following features:

@@

\emph{Type} The possible values for the Type feature include all the basic Python
types (such as int and tuple) as well as a few special values:
\begin{itemize}
    \item IsStatement - Given to all statements, since our terms include both
    statments and expressions but only expressions actually have types.
    \item Unknown - Given to expressions that are never evaluated in the dynamic
    trace.
    \item Multiple - Given to expressions which are evaluated multiple times in
    a trace and do not always have the same type.
\end{itemize}

@@

\emph{Syntactic form} The first feature we consider is what kind of node the term
is in the AST. For statements this has values like Return or Import; for expressions
Variable or Application.

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={Top-1,Top-3},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=withEverything, col sep=comma]{fault-localization.csv};
    \addplot table[x=scoreName, y=baseline, col sep=comma]{fault-localization.csv};
\legend{Our best model,Python interpreter}
\end{axis}
\end{tikzpicture}
\caption{Based on a decision tree on the largest
dataset (272534 pairs). TODO: also train for Top-2}
\label{fig:full-dataset-acc}
\end{figure}

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={Top-1,Top-3},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=all, col sep=comma]{removing-features.csv};
    \addplot table[x=scoreName, y=noTypes, col sep=comma]{removing-features.csv};
    \addplot table[x=scoreName, y=noContext, col sep=comma]{removing-features.csv};
    \addplot table[x=scoreName, y=noSlice, col sep=comma]{removing-features.csv};
\legend{All Features,No Types, No Context, No Slicing}
\end{axis}
\end{tikzpicture}
\caption{Accuracy when features are removed, based on decision trees on a random subset of 20000 pairs.
TODO: train better model and/or
train on greater fraction of the data? TODO: Top-2. NOTE: removing slicing makes basically no difference...}
\label{fig:removing-features}
\end{figure}

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={Top-1,Top-3},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=All, col sep=comma]{defect-categories.csv};
    \addplot table[x=scoreName, y=TypeError, col sep=comma]{defect-categories.csv};
    \addplot table[x=scoreName, y=AttributeError, col sep=comma]{defect-categories.csv};
\legend{All Errors,TypeError,AttributeError}
\end{axis}
\end{tikzpicture}
\caption{Accuracy when using only a subset of the data that all has the same crash. TODO:
get separate baselines?}
\label{fig:defect-categories}
\end{figure}

@@

\begin{figure}
\foreach \method in {single, complete, average}
{
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Threshold,
      ylabel=\# of clusters,
    %   ytick distance=2,
      title=Linkage: \method
    ]
    \addplot+[mark=none] table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
\end{axis}
\end{tikzpicture}
}
\caption{The number of clusters when clustering a baseline of twenty dissimilar programs from
  Rosetta Code. We give no plot for the baseline of 5000 identical programs, as they were
  consistently placed in just 1 cluster.}
\label{fig:diversity-baseline}
\end{figure}

\begin{figure}
\foreach \method in {single, complete, average}
{
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Threshold,
      ylabel=\# of clusters,
    %   ytick distance=1000,
      title=Linkage: \method
    ]
    \foreach \language in {ocaml, python}
    {
      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
    }
\legend{OCaml, Python}
\end{axis}
\end{tikzpicture}
}
\caption{Number of clusters when data is clustered by single, complete, and average linkage.
  A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
  into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
  of a link is its distance, z-scored against other nearby links.}
\label{fig:diversity}
\end{figure}

>>>>

We follow the overall strategy of \cite{learning-to-blame} and represent each
buggy program by a Bag of Abstracted Terms. In our setting, a `term' means either
a statment or an expression.

For each term in the program we compute static, dynamic,
and contextual features. While it is easy to read some features off of the surface
syntax, we expect that some features which are only accessible dynamically will also
be useful, for example the type of each expression. Since programs are highly
structured and expressions gain meaning in relation to their surroundings,
we also expect contextual features to be useful.

\emph{Syntactic form} We expect certain kinds of terms to be more prone to bugs
than others - for example, students might have more trouble with loop conditions
than with simple assignments. Thus the first feature we consider is what kind of node the term
is in the AST. For statements this has values like Return or Import; for expressions
Variable or Application.

@@

\emph{Type} An expression's type should be a very useful feature. Not only are
incompatible types a sign of an impending crash, but some types may be inherently
suspect whenever they appear, e.g. there are very few reasons to have a variable
of type NoneType. Our possible values for the Type feature include all the basic Python
types (such as int and tuple) as well as a few special values:
\begin{itemize}
    \item IsStatement - Given to all statements, since our terms include both
    statments and expressions but only expressions actually have types.
    \item Unknown - Given to expressions that are never evaluated in the dynamic
    trace.
    \item Multiple - Given to expressions which are evaluated multiple times in
    a trace and do not always have the same type.
\end{itemize}

@@

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={Baseline,Top-1,Top-3},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{fault-localization.csv};
\end{axis}
\end{tikzpicture}
\caption{Baseline is the normal Python interpreter. Other bars are from decision
trees on the largest dataset (272534 pairs). TODO: also train for Top-2}
\label{fig:full-dataset-acc}
\end{figure}

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={noContext, noTypes, all},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2.csv};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when features are removed, based on decision trees on a random subset of 20000 pairs.
TODO: Top-2}
\label{fig:removing-features}
\end{figure}

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={All, TypeError, AttributeError},
    xtick=data,
    ymin=0,
    ylabel=Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2.csv};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when the model is trained and tested on only the data exhibiting
the error on the x-axis.}
\label{fig:defect-categories}
\end{figure}

@@

% \begin{figure}
% \foreach \method in {single, complete, average}
% {
% \begin{tikzpicture}
%   \begin{axis}[
%       xlabel=Threshold,
%       ylabel=\# of clusters,
%     %   ytick distance=2,
%       title=Linkage: \method
%     ]
%     \addplot+[mark=none] table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
% \end{axis}
% \end{tikzpicture}
% }
% \caption{The number of clusters when clustering a baseline of twenty dissimilar programs from
%   Rosetta Code. We give no plot for the baseline of 5000 identical programs, as they were
%   consistently placed in just 1 cluster.}
% \label{fig:diversity-baseline}
% \end{figure}

% \begin{figure}
% \foreach \method in {single, complete, average}
% {
% \begin{tikzpicture}
%   \begin{axis}[
%       xlabel=Threshold,
%       ylabel=\# of clusters,
%     %   ytick distance=1000,
%       title=Linkage: \method
%     ]
%     \foreach \language in {ocaml, python}
%     {
%       \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
%     }
% \legend{OCaml, Python}
% \end{axis}
% \end{tikzpicture}
% }
% \caption{Number of clusters when data is clustered by single, complete, and average linkage.
%   A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
%   into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
%   of a link is its distance, z-scored against other nearby links.}
% \label{fig:diversity}
% \end{figure}

\begin{figure}
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Clustering threshold parameter,
      ylabel=\# of naturally occuring program clusters,
      xmax=1.25,
      legend style={at={(0.5,0.25)},anchor=north}
    %   ytick distance=1000,
    %   title=Linkage: single
    ]
    \foreach \language in {ocaml, python}
    {
      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_single, col sep=comma]{cluster_counts.csv};
    }
\legend{NATE OCaml data, Our Python data}
\end{axis}
\end{tikzpicture}
\caption{Number of clusters when data is clustered by single linkage.
  A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
  into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
  of a link is its distance, z-scored against other nearby links. TODO: find a way to include
  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.}
\label{fig:diversity}
\end{figure}
